Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms

### 2019-07-26 manuscript/main.tex
Update text and figures

\documentclass{article}
\usepackage{natbib}
\usepackage{subfiles}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{graphicx}
\graphicspath{{../Figures/}}

@@

Microbial admixture could arise through a
number of ecological mechanisms illustrated in conceptual diagram 1. For
example, an environmental condition such as pH could contribute to a high rate
of ecosystem function and select for a taxon even if that taxon is unrelated to
function. An association could also arise if an organism disperses from a
low-rate ecosystem to an ecosystem with a high rate of ecosystem function
through source-sink dynamics. In fact, if the rate of
ecosystem function is high in one set of ecosystems, function will be
associated with any taxa that are abundant in those ecosystems.
In order to understand the biological basis of ecosystem function, we need to
separate the close association between environmental variation, community
composition, and ecosystem function in order to identify what aspects of
microbial biodiversity determine the rate of ecosystem function.

@@

In this paper we propose integrating microbial biodiversity-ecosystem function
research with ideas and techniques from population genetics to advance our
understanding of the role of microbial communities in the functioning of
ecosystems. Perhaps the most important part of that discussion is to define the
question carefully. Microbial structure-function research should focus on answering the question: what is the
unique contribution of the microbial community to ecosystem function independent
of the environment? This is of
great interest for two reasons. First, because it addresses what is the singular
contribution of biodiversity to the functioning of ecosystems. Second, on a
practical level, because answering that question will be essential to
incorporate a microbial component into ecosystem models. To address that
question, we demonstrate that one ecosystem function, high-affinity methane
oxidation, is significantly association with variation in microbial community
composition. We also show that three components of population stratification
contribute significantly to variation in ecosystem function. Finally, we
demonstrate one approach to identifying the component of microbial
community that contributes to ecosystem function use an association test that
controls for these three
covariates and find four microbial taxa that are significantly associated with
ecosytsem function.

@@

All statistical analyses were performed using the \texttt{R} statistical
environment. 
We first applied traditional tests of microbial biodiversity-ecosystem function
relationships. To do this, we fit ordinary least squares regression models with
the rate of ecosystem function as the response with one predictor for each
model. These predictors include functional gene and transcript abundance,
diversity measures including richness, Shannon's diversity index, and Simpson's
diversity index, measures of microbial community composition using the first
several principal component axes from a principal component
analysis (PCA) chosen based on visual inspection of a screeplot incpluding a PCA for
abundance data and a PCA for presence/absence data, and several 
environmental variables.

Next, we used a statistical approach common in population genetics genome-wide
association studies. We fit variance component mixed models from the
\texttt{varComp} package to test whether variation in microbial
community composition expains a significant portion of the variation in
ecosystem function \citep{qu2013}. First, we fit an intercept-only model with
each random effect in all combinations. Variance components included deographic location was coded as a factor, environmental
similarity from euclidean distance of four edaphic variables including total
nitrogen, total carbon, moisture content, and bulk density, and
community similarity calculated from Bray-Curtis dissimilarity for abundance
data and the Jaccard index for presence/absance data. We tested the significance
of these components using a linear score test by comparing nested models with and
without each variance component \citep{qu2013}. We then estimated
the percent of variation in ecosystem function explained by each component 
including standard errors on that estimate 
using the \texttt{h2G} function from the
\texttt{gap} package. We then tested whether microbial community composition explains a
significant portion of the variation in ecosystem function after accounting for
variation in the environment and accounting for the spatial distance between
samples. We did this by fitting an intercept-only model with variance components
for environmental similarity from euclidean distance of a matrix of
environmental covariates and spatial similarity
also using euclidean distance from the latitude and longitude coordinates where
the samples were collected. We then used a linear score test to compare
nested models with and without the microbial community variance component to
determine whether community composition adds any variance explained after
accounting for spatial and environmental covariates. We then estimated the
variance explained by each component (environment, geography, and community
composition) with standard errors of the estimates using the \texttt{h2GE}
function.

@@

\includegraphics[width=0.5\textwidth]{lowk_pmoa}
\includegraphics[width=0.5\textwidth]{lowk_pc1}
\includegraphics[width=0.5\textwidth]{lowk_rich}
\includegraphics[width=0.5\textwidth]{lowk_shan}

All three variance components are significant individually. Community
significant in the community + environment model and geography
significant in the geography + environment model. Table \ref{tab:pvalues}

@@

Including different covariates removes different numbers of
significant taxa. Some covariates add taxa that were not previously
identified as significant. Table \ref{tab:n_taxa}

@@

After controlling for geographic proximity, environmental similarity, and
similarity in community composition, four taxa were significantly correlated
with the rate of methane oxidation. None of these four taxa are known methane
oxidizers or methane producers. Table  \ref{tab:sig_taxa}

@@

In this paper, we argue that the traditional approach to microbial BEF research
ignores important understandings about the nature of ecological processes in
determining the distribution of microorganisms. The marker gene hypothesis has
proven unsuccesful in predicting the rates of ecosystem functions from the
abundance of functional genes and in our example study was not significant.
Similarly, diversity metrics from the entire Prokaryotic community were not
explanatory. These results agree with recent reviews of this literature and
demonstrate that these approaches are not a fruitful avenue for elucidating
microbial structure-function connections.

>>>>

\documentclass{article}
\usepackage{natbib}
\usepackage{subfiles}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{graphicx}
\graphicspath{{../figures/}}

@@

Microbial admixture could arise through a
number of ecological mechanisms illustrated in the conceptual diagram in Figure 1. For
example, an environmental condition such as pH could contribute to a high rate
of ecosystem function and select for a taxon even if that taxon is unrelated to
function. An association could also arise if an organism disperses from a
low-rate ecosystem to an ecosystem with a high rate of ecosystem function
through source-sink dynamics. In fact, if the rate of
ecosystem function is high in one set of ecosystems, function will be
associated with any taxa that are abundant in those ecosystems.
In order to understand the biological basis of ecosystem function, we need to
separate the close association between environmental variation, community
composition, and ecosystem function in order to identify what aspects of
microbial biodiversity determine the rate of ecosystem function.

@@

In this paper we propose integrating microbial biodiversity-ecosystem function
research with ideas and techniques from population genetics to advance our
understanding of the role of microbial communities in the functioning of
ecosystems. Perhaps the most important part of that discussion is to define the
question carefully. Microbial structure-function research should focus on answering the question: what is the
unique contribution of the microbial community to ecosystem function independent
of the environment? This is of
great interest for two reasons. First, because it addresses what is the singular
contribution of biodiversity to the functioning of ecosystems. Second, on a
practical level, because answering that question will be essential to
incorporate a microbial component into ecosystem models. To address that
question, we demonstrate that one ecosystem function, high-affinity methane
oxidation, is significantly association with variation in microbial community
composition. We also show that three components of population stratification
contribute significantly to variation in ecosystem function. Finally, we
demonstrate one approach to identifying the component of microbial
community that contributes to ecosystem function using an association test that
controls for these three
covariates and find four microbial taxa that are significantly associated with
ecosytsem function.

@@

All statistical analyses were performed using the \texttt{R} statistical
environment. 
We first applied traditional tests of microbial biodiversity-ecosystem function
relationships. To do this, we fit ordinary least squares regression models with
the rate of ecosystem function as the response with one predictor for each
model. These predictors include functional gene abundance,
diversity measures including richness, Shannon's diversity index, and Simpson's
diversity index, measures of microbial community composition using the first
several principal component axes from a principal component
analysis (PCA) chosen based on visual inspection of a screeplot incpluding a PCA for
abundance data and a PCA for presence/absence data, and several 
environmental variables.

Next, we used a statistical approach common in population genetics genome-wide
association studies. We fit variance component mixed models from the
\texttt{varComp} package to test whether variation in microbial
community composition expains a significant portion of the variation in
ecosystem function \citep{qu2013}. First, we fit an intercept-only model with
each random effect in all combinations. Variance components included geographic location was coded as a factor, environmental
similarity from euclidean distance of four edaphic variables including total
nitrogen, total carbon, moisture content, and bulk density, and
community similarity calculated from Bray-Curtis dissimilarity for abundance
data and the Jaccard index for presence/absance data. We tested the significance
of these components using a linear score test by comparing nested models with and
without each variance component \citep{qu2013}. We then estimated
the percent of variation in ecosystem function explained by each component 
including standard errors on that estimate 
using the \texttt{h2G} function from the
\texttt{gap} package. We then tested whether microbial community composition explains a
significant portion of the variation in ecosystem function after accounting for
variation in the environment and accounting for the spatial distance between
samples. We did this by fitting an intercept-only model with variance components
for environmental similarity from euclidean distance of a matrix of
environmental covariates and spatial similarity
also using euclidean distance from the latitude and longitude coordinates where
the samples were collected. We then used a linear score test to compare
nested models with and without the microbial community variance component to
determine whether community composition adds any variance explained after
accounting for spatial and environmental covariates. We then estimated the
variance explained by each component (environment, geography, and community
composition) with standard errors of the estimates using the \texttt{h2GE}
function.

@@

Methane oxidation rate was significantly correlated with the first principal
component from a PCA of the
community matrix of ASVs inferred from 16S rRNA genes (Figure
\ref{fig:corrs}, Table 1). Other typical
measurements of community structure including functional gene abundance for the
gene encoding the final step in the methanotrophy pathway, richness, and Shannon
diversity were not significantly correlated with methane oxidation rate.

\subfile{../tables/trad_cor_table}

\begin{figure}
\includegraphics[width=0.5\textwidth]{lowk_pmoa}
\includegraphics[width=0.5\textwidth]{lowk_pc1}
\includegraphics[width=0.5\textwidth]{lowk_rich}
\includegraphics[width=0.5\textwidth]{lowk_shan}
\caption{\label{fig:corrs}
  \textbf{Methane oxidation rate is correlated with community composition, but not
functional gene abundance or diversity}.
  Correlations between community attributes and ecosystem function. A) Abundance
  of the functional gene \textit{pmoA}, B) community composition based on
  principal component 1 of a PCA, C) richness, and D) Shannon's diversity index. n = 44 except
for A where n = 42, because two samples failed to amplify during quantitative
PCR.}
\end{figure}

Three variance components explained significant portion of the variation in methane
oxidation when tested individually including community similaity from
Bray-Curtis distance, environmental similarity, and geographic similarity. When
combined, most of these terms became not significant with two exceptions. In the
community + environment model, community was still significant and in the
geography + environment model, geography was still significant (Table 2).

@@

There were 460 taxa that were significantly correlated with methane oxidation
rate. After including each covariate, fewer taxa were significant. All variance
components removed some significant taxa and several of the models added taxa 
that were not previously
identified as significant. Controlling for community similarity removed all
significant taxa and controlling for geography and community removed all
significant taxa. After controlling for all three covariates, only four taxa
remained significant (Table 3).

@@

After controlling for geographic proximity, environmental similarity, and
similarity in community composition, four taxa were significantly correlated
with the rate of methane oxidation. None of these four taxa are known methane
oxidizers or methane producers (needs citations) (Table 4).

@@

In this paper, we argue that the traditional approach to microbial BEF research
ignores important understandings about the nature of ecological processes in
determining the distribution of microorganisms. The marker gene hypothesis has
proven unsuccesful in predicting the rates of ecosystem functions from the
abundance of functional genes and in our example study was not a good predictor
of ecosystem function.
Similarly, diversity metrics from the entire Prokaryotic community were not
explanatory. These results agree with recent reviews of this literature and
demonstrate that these approaches are not a fruitful avenue for elucidating
microbial structure-function connections.
